---
title: "Custom metric sets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom_metric_sets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=F}
library(lidR)
library(lidRmetrics)
```

The lidRmetrics package is designed to provide a collection of functions for calculating point cloud metrics. Each function is specialized in computing metrics that characterize a specific aspect of the point cloud structure. This modular approach allows users to select and combine functions to create a custom set of metrics tailored to their specific needs.

The package includes three sets of predefined metrics, but users are encouraged to develop their own. The following examples demonstrate how to do this.

We'll begin with a simple example by combining metrics_basic() and metrics_percentiles():

```{r}
my_metric_set <- function(z) {
  
  m_basic   <- lidRmetrics::metrics_basic(z=z)
  m_prctls  <- lidRmetrics::metrics_percentiles(z=z)
  
  m <- c(m_basic, m_prctls)
  
  return(m)
}
```

It's important to note that each function is explicitly called with the lidRmetrics package prefix. This practice is crucial when processing data on multiple cores using the future package.

Let's test the function and examine the outputs:

```{r}
LASfile <- system.file("extdata", "Megaplot.laz", package="lidR")
las <- readLAS(LASfile, select = "*", filter = "-keep_random_fraction 0.5")

my_metrics <- pixel_metrics(las, ~my_metric_set(Z), res = 20)

head(as.data.frame(my_metrics),1)
```


Now, let's consider a more complex scenario. Suppose we want to calculate metrics_basic() and metrics_percentiles(), but for each metric, we want to consider two different scenarios: with and without a height cutoff of 2 meters (zmin). We can easily adapt the previous function to calculate both variants of each metric:

```{r}

my_metric_set <- function(z) {
  
  # set the height threshold
  zmin <- 2
  
  # define the text that will be appended to metric names calculated with height threshold. This is to avoid duplicated metric names.
  zmin_label <- paste0(".above",zmin)
  
  # metrics_basic without height threshold
  m_basic   <- lidRmetrics::metrics_basic(z=z)
  
  # metrics_basic with height threshold. To make metric names unique, `zmin_label` is appended.
  m_basic_2   <- lidRmetrics::metrics_basic(z=z, zmin=zmin)
  names(m_basic_2) <- paste0(names(m_basic_2),zmin_label)
  
  # the same for metrics_percentiles
  m_prctls  <- lidRmetrics::metrics_percentiles(z=z)
  m_prctls_2  <- lidRmetrics::metrics_percentiles(z=z, zmin=zmin)
  names(m_prctls_2) <- paste0(names(m_prctls_2),zmin_label)
  
  m <- c(m_basic, m_prctls, m_basic_2, m_prctls_2)
  return(m)
}

```

The output of the function above will look as follows:
```{r, echo=F}
LASfile <- system.file("extdata", "Megaplot.laz", package="lidR")
las <- readLAS(LASfile, select = "*", filter = "-keep_random_fraction 0.5")

my_metrics <- pixel_metrics(las, ~my_metric_set(Z), res = 20)
```

```{r}
head(as.data.frame(my_metrics),1)
```


Finally, metrics can be calculated for different echo types (e.g., for first returns only, for all returns, etc.). This can be achieved using the built-in functionality of the lidR package:

```{r}
my_metrics <- pixel_metrics(las, ~my_metric_set(Z), by_echo=c("first", "all"))
```

The output will include the results of metrics_basic() and metrics_percentiles(), calculated for the following scenarios: all returns with no height threshold, all returns above 2 meters, first returns only with no height threshold, and finally, first returns above 2 meters. Since point cloud subsetting is performed during metric calculation, it is important to avoid applying any filtering when loading the point cloud data or creating the catalog.

```{r}
head(as.data.frame(my_metrics),1)
```
